### 2019-03-02 paper/sex_selection_spacing-ver2-new-intro.tex
Minor edits

% [What do we know about use of sex selection?]
Over the last three decades, India has seen dramatic increases in the males-to-females 
ratio at birth as access to prenatal sex determination expanded 
\citep{das_gupta97,Sudha1999,Arnold2002,retherford03b,jha06,Guilmoto2012,Portner2015b}.
Also, fertility has declined substantially, to the point where it is now
at, or even below, replacement in some areas 
\citep{Guilmoto2013,Dharmalingam2014,International-Institute-for-Population-Sciences-IIPS2017}.
This decline in fertility, combined with a strong preference for sons, has intensified 
the pressure to use sex selection \citep{Guilmoto2009a,Bongaarts2013,Jayachandran2017}.%
\footnote{
In some instances, however, the fertility decline appears to have created a stronger 
aversion to daughters rather than a stronger preference for son \citet{DiamondSmith2008}.
}
Within families, the strongest predictor of the use of sex selection at a given parity is 
the sex composition of previous children \citep{retherford03b,jha06,abrevaya09,Portner2015b}.
For families without a son, the higher the parity is, the higher is the probability
of having a son as the next birth.
The use of sex selection also increases with socioeconomic status, especially 
education, and is more prevalent in cities than in rural areas, both possibly because 
of lower desired fertility \citep{retherford03b,jha06,Portner2015b}.

@@

Despite the potential important impact of birth spacing on outcomes for both
mother and children, there has so far been no examination of what effect access to
sex selection has had on spacing.

@@

We know that health outcomes for girls appear to improve in the presence of sex 
selection \citep{Lin2014,Hu2015}.
It is possible that these improvements are an unintended side-effect of the longer spacing 
that arises from sex-selective abortions, rather than because the smaller number of girls 
makes them more valued as is often assumed.
A first step

% [Results]
There are three main results.
First, there has been a general increase in the length of spacing between births
over the four decades covered by the data.
Second, the most substantial increase in spacing is for the women who
are most likely to use sex selection.
Specifically, among the best-educated women, those with no boys now has 
significantly longer spacing---and a more male-biased sex ratio---than 
similar women with boys.
Third, women with no education still follow the standard pattern of
short spacing when they have girls and little evidence of the use of sex
selection.
In other words, these women adhere to a strategy where they achieve a
son through higher fertility rather than the use of sex selection.

>>>>

% [What do we know about use of sex selection?]
Over the last three decades, India has seen dramatic increases in the males-to-females 
ratio at birth as access to prenatal sex determination expanded 
\citep{das_gupta97,Sudha1999,Arnold2002,retherford03b,jha06,Guilmoto2012,Portner2015b}.
Also, fertility has declined substantially, to the point where it is now
at, or even below, replacement in some areas 
\citep{Guilmoto2013,Dharmalingam2014,International-Institute-for-Population-Sciences-IIPS2017}.
This decline in fertility, combined with a strong preference for sons, has intensified 
the pressure to use sex selection 
\citep{das_gupta97,Guilmoto2009a,Bongaarts2013,Jayachandran2017}.%
\footnote{
In some instances, however, the fertility decline appears to have created a stronger 
aversion to daughters rather than a stronger preference for son \citet{DiamondSmith2008}.
}
Within families, the strongest predictor of the use of sex selection at a given parity is 
the sex composition of previous children \citep{retherford03b,jha06,abrevaya09,Portner2015b}.
For families without a son, the higher the parity is, the higher is the probability
of having a son as the next birth.
The use of sex selection also increases with socioeconomic status, especially 
education, and is more prevalent in cities than in rural areas, both possibly because 
of lower desired fertility \citep{retherford03b,jha06,Portner2015b}.

@@

Despite the increasing use of sex selection, the extensive use of spacing to measure son 
preference, and the important impacts of spacing on outcomes for both mother and children  
there has so far been no examination of what effect access to sex selection has had on 
spacing.

@@

[move to conclusion?]
We know that health outcomes for girls appear to improve in the presence of sex 
selection \citep{Lin2014,Hu2015}.
It is possible that these improvements are an unintended side-effect of the longer spacing 
that arises from sex-selective abortions, rather than because the smaller number of girls 
makes them more valued as is often assumed.
A first step

% [Results - need updating]
There are three main results.
First, there has been a general increase in the length of spacing between births
over the four decades covered by the data.
Second, the most substantial increase in spacing is for the women who
are most likely to use sex selection.
Specifically, among the best-educated women, those with no boys now has 
significantly longer spacing---and a more male-biased sex ratio---than 
similar women with boys.
Third, women with no education still follow the standard pattern of
short spacing when they have girls and little evidence of the use of sex
selection.
In other words, these women adhere to a strategy where they achieve a
son through higher fertility rather than the use of sex selection.
